```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/home/jkincade/Methyl/4mo/BAMs/")
folder = "/home/jkincade/Methyl/4mo/TI/Results/"
date = format(Sys.Date(), '%m-%d-%Y')
```


# Installing Required Packages 
## It is only necessary to do this the first time running this code

```{r}
# ## Installing required packages
# install.packages("foreach")
# install.packages("tidyr")
# install.packages("BiocManager")
# BiocManager::install("genomation")
# BiocManager::install("methylKit")
# BiocManager::install("TxDb.Btaurus.UCSC.bosTau9.refGene")
# install.packages("vegan")
# install.packages("ggbio")
# install.packagges("pheatmap")
```

# Loading the packages
## This will be done each time you open RStudio and aim to run this code

```{r}
library(foreach)
library(BiocManager)
library(genomation)
library(methylKit)
library(parallel)
library(TxDb.Btaurus.UCSC.bosTau9.refGene)
library(ggbio)
library(pheatmap)
library(ggpubr)
library(vegan)
```

# 'CpG' can be replaced with either CHG or CHH to generate CHG or CHH sites in addition to CpG sites; only CpG sites are shown here. 


# Aligning .bam files to the genome to generate bismark files

```{r}
# Create a list of .bam files
bam.list <- list("2103_TI_R1_val_1_bismark_bt2_pe.bam", "2106_TI_R1_val_1_bismark_bt2_pe.bam", "2132_control_R1_val_1_bismark_bt2_pe.bam", "2147_control_R1_val_1_bismark_bt2_pe.bam", "2149_control_R1_val_1_bismark_bt2_pe.bam", "2156_control_R1_val_1_bismark_bt2_pe.bam", "2163_control_TI_R1_val_1_bismark_bt2_pe.bam", "2169_TI_R1_val_1_bismark_bt2_pe.bam", "2183_TI_R1_val_1_bismark_bt2_pe.bam", "2186_TI_R1_val_1_bismark_bt2_pe.bam")



# Pair the files to more accessible sample IDs
ids <- list("2103.TI", "2106.TI", "2132.C", "2147.C", "2149.C", "2156.C", "2163.C", "2169.TI", "2186.TI", "2183.TI")



# Indicating treatment groups
treatment.list <- c(1, 1, 0, 0, 0, 0, 0, 1, 1, 1)

# Alignment of the .bam files

foreach(i=1:10) %dopar% {
  parallel_CpG_call <- processBismarkAln(
    location=bam.list[i],
    sample.id=ids[i],
    assembly = "bosTau9",
    save.folder = "/home/jkincade/Methyl/4mo/Call-Lists/",
    save.context = "CpG", read.context = "CpG",
    treatment = treatment.list[i])
}
```

# Read in the aligned data

```{r}
# Move your working directory to where the bismark files were stored
setwd("/home/jkincade/Methyl/4mo/Call-Lists/")



# Create a call list 

CpG.Call.List <- list.files(pattern = ".*\\CpG.txt$")
CpG.Call.List # Take note of the sample id order here!

CpG.Call.List <- as.list(CpG.Call.List)

# Read in the bismark files. Make sure that this list reflects the order of the CpG.Call.List in both the ids and treatment. 
# For treatment = c()     0 = control, 1 = treatment 


CpG.Read <- methRead(CpG.Call.List, sample.id = list("2103.TI", "2106.TI", "2132.C", "2147.C", "2149.C", "2156.C", "2163.C", "2169.TI", "2183.TI", "2186.TI"),
                     assembly = "bosTau9", 
                     treatment = c(1, 1, 0, 0, 0, 0, 0, 1, 1, 1), 
                     context = "CpG")
```


# Generating descriptive statistics

```{r}
# Methylation statistics
## Large peaks on either end of the histogram are normal
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylStats_2103.TI.pdf')
getMethylationStats(CpG.Read[[1]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylStats_2106.TI.pdf')
getMethylationStats(CpG.Read[[2]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylStats_2132.C.pdf')
getMethylationStats(CpG.Read[[3]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylStats_2147.C.pdf')
getMethylationStats(CpG.Read[[4]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylStats_2149.C.pdf')
getMethylationStats(CpG.Read[[5]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylStats_2156.C.pdf')
getMethylationStats(CpG.Read[[6]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylStats_2163.C.pdf')
getMethylationStats(CpG.Read[[7]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylStats_2169.TI.pdf')
getMethylationStats(CpG.Read[[8]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylStats_2183.TI.pdf')
getMethylationStats(CpG.Read[[9]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylStats_2186.TI.pdf')
getMethylationStats(CpG.Read[[10]], plot = TRUE, both.strands = FALSE)
dev.off()



# Coverage statistics 
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylCoverage_2103.TI.pdf')
getCoverageStats(CpG.Read[[1]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylCoverage_2106.TI.pdf')
getCoverageStats(CpG.Read[[2]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylCoverage_2132.C.pdf')
getCoverageStats(CpG.Read[[3]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylCoverage_2147.C.pdf')
getCoverageStats(CpG.Read[[4]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylCoverage_2149.C.pdf')
getCoverageStats(CpG.Read[[5]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylCoverage_2156.C.pdf')
getCoverageStats(CpG.Read[[6]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylCoverage_2163.C.pdf')
getCoverageStats(CpG.Read[[7]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylCoverage_2169.TI.pdf')
getCoverageStats(CpG.Read[[8]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylCoverage_2183.TI.pdf')
getCoverageStats(CpG.Read[[9]], plot = TRUE, both.strands = FALSE)
dev.off()
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_MethylCoverage_2186.TI.pdf')
getCoverageStats(CpG.Read[[10]], plot = TRUE, both.strands = FALSE)
dev.off()



```
# Filtering and normalization 
```{r}
# Filtering by coverage

filtered.CpG.Read <- filterByCoverage(CpG.Read, lo.count = 10, lo.perc = NULL,hi.count = NULL, hi.perc = 99.9)



# Normalization
## method = "median" OR "mean" 

normal.filtered.CpG.Read <- normalizeCoverage(filtered.CpG.Read, method = "median")
united.CpG <- methylKit::unite(normal.filtered.CpG.Read, destrand = TRUE) # destrand = TRUE can improve coverage for CpG, but is not recommended for CHG or CHH



# Filtering by standard deviation (graphs compare the matrix before and after filtering by standard deviation)

percent.meth.CpG <- percMethylation(united.CpG)

CpG.matrix <- matrixStats::rowSds(percent.meth.CpG)

pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_PercentMethylation_Before_Filtering_CpG.pdf')
hist(CpG.matrix, col = 'cornflowerblue', xlab = 'Std. Dev. per CpG', main="Before Filtering", breaks = 100)
dev.off()

pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_PercentMethylation_After_Filtering_CpG.pdf')
par(mar = c(4.1, 4.4, 4.1, 1.9), xaxs="i", yaxs="i")
hist(
  CpG.matrix[CpG.matrix > 2],
  col = 'cornflowerblue',
  xlab = 'Std. Dev. per CpG',
  breaks = 100,
  main = "After Filtering"
)
dev.off()

# See what percentage of sites were retained after filtering by standard deviation
print("CpGs before filtering = ")
nrow(united.CpG)
before <- nrow(united.CpG)
stdev.filtered.CpG <- united.CpG[CpG.matrix > 2]
print("CpGs after filtering = ")
nrow(stdev.filtered.CpG)
after <- nrow(stdev.filtered.CpG)
print("Percent CpGs Retained = ")
round(after/before*100,2)



```
# Correlation Plot (be patient)

```{r}
# CpG
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_Correlation_CpG.pdf')
par(mar = c(4.1, 4.4, 4.1, 1.9), xaxs="i", yaxs="i")
getCorrelation(united.CpG, plot = TRUE)
dev.off()

```


# Clustering and PCA plots 
```{r}
# These plots use ALL of the CpG sites after filtering but before differential analysis 

pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_TIvsC_Clustering.pdf')
clusterSamples(united.CpG, dist = "correlation", method = "ward.D2", plot = TRUE)
dev.off()


pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_PCA_Plot_TIvsC.pdf') 
CpGPCA = PCASamples(united.CpG, obj.return = TRUE)
dev.off()

summary(CpGPCA)

pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_PCA_Screeplot_TIvsC.pdf')
PCASamples(united.CpG, screeplot = TRUE)
dev.off()


```
# Removing 'chrUNs' (chromosome unknown)

```{r}
rowlist <- list()
L <- 1
for(i in 1:nrow(united.CpG)){
  chr <- united.CpG$chr [i]
  if(grepl("chrUn",chr)){
    rowlist [[L]] <- i
    L <- L + 1
  }
}
#rowlist
rw.length <- length(rowlist)
rw.length.minus <- rowlist[[1]] - 1
rowlist[[rw.length]]
united.CpG <- united.CpG[-c(rowlist[[1]]:rowlist[[rw.length]]),]
```


# Calculating differentially methylated sites (DMS)

```{r}
CpG.Diff <- calculateDiffMeth(united.CpG, mc.cores = 30)

# "MN" (McCullagh and Nedler) applies a basic overdispersion correction
CpG.Diff1 <- calculateDiffMeth(united.CpG, overdispersion = "MN", test = "Chisq", mc.cores = 30)


# CpG Stats
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/CpG_Stats_TIvsC.pdf')
methSeg(
CpG.Diff1,
diagnostic.plot = TRUE,
join.neighbours = FALSE)
dev.off()
```


# Determining whether you should correct for overdispersion or not
```{r}
# These functions return the number of differentially methylated sites with varying values for 'meth.diff' when the data IS NOT corrected for overdispersion 

a = dim(getMethylDiff(CpG.Diff, difference = 10, qvalue = 0.01))[1]
print(paste("Number of sites with difference 10 for data that corrected for overdispersion is", a))
b = dim(getMethylDiff(CpG.Diff, difference = 15, qvalue = 0.01))[1]
print(paste("Number of sites with difference 15 for data that corrected for overdispersion is", b))
c = dim(getMethylDiff(CpG.Diff, difference = 20, qvalue = 0.01))[1]
print(paste("Number of sites with difference 20 for data that corrected for overdispersion is", c))
d = dim(getMethylDiff(CpG.Diff, difference = 25, qvalue = 0.01))[1]
print(paste("Number of sites with difference 25 for data that corrected for overdispersion is", d))
e = dim(getMethylDiff(CpG.Diff, difference = 30, qvalue = 0.01))[1]
print(paste("Number of sites with difference 30 for data that corrected for overdispersion is", e))
f = dim(getMethylDiff(CpG.Diff, difference = 35, qvalue = 0.01))[1]
print(paste("Number of sites with difference 35 for data that corrected for overdispersion is", f))
g = dim(getMethylDiff(CpG.Diff, difference = 40, qvalue = 0.01))[1]
print(paste("Number of sites with difference 40 for data that corrected for overdispersion is", g))
h = dim(getMethylDiff(CpG.Diff, difference = 50, qvalue = 0.01))[1]
print(paste("Number of sites with difference 50 for data that corrected for overdispersion is", h))
i = dim(getMethylDiff(CpG.Diff, difference = 75, qvalue = 0.01))[1]
print(paste("Number of sites with difference 75 for data that corrected for overdispersion is", i))

# These functions return the number of differntially methylated sites with varying values for 'meth.diff' when the data IS corrected for overdispersion 

a1 = dim(getMethylDiff(CpG.Diff1, difference = 10, qvalue = 0.01))[1]
print(paste("Number of sites with difference 10 for data is", a1))
b1 = dim(getMethylDiff(CpG.Diff1, difference = 15, qvalue = 0.01))[1]
print(paste("Number of sites with difference 15 for data is", b1))
c1 = dim(getMethylDiff(CpG.Diff1, difference = 20, qvalue = 0.01))[1]
print(paste("Number of sites with difference 20 for data is", c1))
d1 = dim(getMethylDiff(CpG.Diff1, difference = 25, qvalue = 0.01))[1]
print(paste("Number of sites with difference 25 for data is", d1))
e1 = dim(getMethylDiff(CpG.Diff1, difference = 30, qvalue = 0.01))[1]
print(paste("Number of sites with difference 30 for data is", e1))
f1 = dim(getMethylDiff(CpG.Diff1, difference = 35, qvalue = 0.01))[1]
print(paste("Number of sites with difference 35 for data is", f1))
g1 = dim(getMethylDiff(CpG.Diff1, difference = 40, qvalue = 0.01))[1]
print(paste("Number of sites with difference 40 for data is", g1))
h1 = dim(getMethylDiff(CpG.Diff1, difference = 50, qvalue = 0.01))[1]
print(paste("Number of sites with difference 50 for data is", h1))
i1 = dim(getMethylDiff(CpG.Diff1, difference = 75, qvalue = 0.01))[1]
print(paste("Number of sites with difference 75 for data is", i1))


library (xtable)
# This creates a table displaying the varying numbers of DMS identified with varying 'meth.diff' WITHOUT dispersion 

diff = c( 10, 15, 20, 25,  30, 35, 40, 50, 75)
DMS = c(a1, b1, c1, d1, e1, f1, g1, h1, i1)
df_DMS = data.frame(diff, DMS)
xtable(df_DMS)

# This creates a table displaying the varying numbers of DMS identified with varying 'meth.diff' WITH dispersion 

diff_c = c( 10, 15, 20, 25, 30, 35, 40, 50, 75)
DMS_c = c(a, b, c, d, e, f, g, h, i)
df_DMS_c = data.frame(diff_c, DMS_c )
xtable(df_DMS_c)




# Bar plot of numbers of DMS with varying meth.diff

## WITHOUT overdispersion correction 
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_Barplot_WITHOUT_overdisp.cor_TIvsC.pdf')
 ggplot(df_DMS, aes(x = diff, y = DMS)) +
   geom_bar(stat = "identity", position = position_dodge()) +
  
   theme_classic() + ylab("Number of DMS + Disp") + xlab("Meth.Diff")+
   scale_color_brewer()+ theme(legend.position = "none")+theme_bw()
 dev.off()
 
 ## WITH overdispersion correction 
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_Barplot_WITH_overdisp.cor_TIvsC.pdf')
 ggplot(df_DMS_c, aes(x = diff, y = DMS_c)) +
   geom_bar(stat = "identity", position = position_dodge()) +
  
   theme_classic() + ylab("Number of DMS + Disp") + xlab("Meth.Diff")+
   scale_color_brewer()+ theme(legend.position = "none")+theme_bw()
 dev.off()
 
# Line graphs of the number of DMS with varying meth.diff

pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_LineGraph_WandWO_overdisp.cor_TIvsC.pdf')
plDMS = ggplot(df_DMS, aes(x = diff, y = DMS)) +
  geom_point(size = 3) +
  
    geom_line(col = "blue")+
 
  ylab("Number of DMS") + xlab("Meth.Diff")+
  geom_vline(xintercept = 25, linetype = "dashed", color = "black", size = 0.5)+
  geom_hline(yintercept = 2325, linetype = "dashed", color = "black", size = 0.5)+
  
  theme(legend.position = "none")+
  theme_classic()+
 theme(axis.text.x=element_text(size=20))+
 theme(axis.text.y=element_text(size=20))
theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))
  scale_color_manual(values = c("Blue Line" = "blue"), name = "Legend Title")


plDMS + geom_line(data = df_DMS_c, aes(x = diff_c, y = DMS_c), color = "red")+ 
geom_point(data = df_DMS_c, aes(x = diff, y =  DMS_c), color = "red", size = 3)+
geom_hline(yintercept = 244, linetype = "dashed", color = "black", size = 0.5)+
   scale_color_manual(values = c("Red Line" = "red"), name = "New Legend") +
   guides(color = guide_legend(title = "Legend Title"))
dev.off()

## The blue line demonstrates the number of DMS at varying meth.diff WITHOUT overdispersion correction 
## The red line demonstrates the number of DMS at varying meth.diff WITH overdispersion correction 
```
# I went forward without correcting for overdispersion, so I renamed the general variable to reflect the method I had chosen

```{r}
CpG.Diff <- calculateDiffMeth(united.CpG, mc.cores = 30)
```

# Get the CpGs from the CpG Diff analysis 
```{r}
## get the hyper methylated bases 
CpG.hyper <- getMethylDiff(CpG.Diff, difference = 25, qvalue = 0.01, type = "hyper")

## get the hypo methylated bases
CpG.hypo <- getMethylDiff(CpG.Diff, difference = 25, qvalue = 0.01, type = "hypo")

## get all of the differentially methylated bases 
CpG.all <-  getMethylDiff(CpG.Diff, difference = 25, qvalue = 0.01)
```




# Graphing DMS per chromosome
```{r}
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_TIvsC_CpG_DMperChr.pdf')
diffMethPerChr(CpG.Diff, plot = TRUE, qvalue.cutoff = 0.01, meth.cutoff = 25)
dev.off()
```
# Circos plots

```{r}
# Get the chromosome lengths for Bos Taurus 
library(TxDb.Btaurus.UCSC.bosTau9.refGene)
chr.len = seqlengths(TxDb.Btaurus.UCSC.bosTau9.refGene)
chr.len = chr.len[grep("_|M|X|Y", names(chr.len), invert = T)]



# Circos plot

ideoDMC <- function(methylDiff.obj, chrom.length, difference = 25, 
                    qvalue = 0.01, circos = FALSE, title = "test", hyper.col = "magenta", 
                    hypo.col = "green") {
  require(methylKit)
  require(GenomicRanges)
  require(ggbio)
  
  # chrom.length
  chrom.length = chr.len
  myIdeo <- GRanges(seqnames = names(chrom.length), ranges = IRanges(start = 1, 
                                                                     width = chrom.length))
  seqlevels(myIdeo) = names(chrom.length)
  seqlengths(myIdeo) = (chrom.length)
  
  
  hypo = getMethylDiff(methylDiff.obj, difference = difference, qvalue = qvalue, 
                        type = "hypo")
  hyper = getMethylDiff(methylDiff.obj, difference = difference, qvalue = qvalue, 
                         type = "hyper")
  
  g.per = as(hyper, "GRanges")
  seqlevels(g.per, force=TRUE) = seqlevels(myIdeo)
  seqlengths(g.per)=(chrom.length)
  
  g.po = as(hypo, "GRanges")
  seqlevels(g.po, force=TRUE) = seqlevels(myIdeo)
  seqlengths(g.po)=(chrom.length)
  
  values(g.po)$id = "hypo"
  values(g.per)$id = "hyper"
  
  if (circos) {
    
    p <- ggplot2::ggplot() + layout_circle(myIdeo, geom = "ideo", fill = "gray70", 
                                  radius = 39, trackWidth = 2)
    
    
    p <- p + layout_circle(c(g.po, g.per), geom = "point", 
                           size = 3, aes(x = midpoint, 
                                         y = meth.diff, color = id), radius = 25, trackWidth = 30) +              
      scale_colour_manual(values = c(hyper.col, hypo.col))
    p + layout_circle(myIdeo, geom = "text", aes(label = seqnames), 
                      vjust = 0, radius = 55, trackWidth = 7) + labs(title = title)
    
  } else {
    
    p <- ggplot2::ggplot() + layout_karyogram(myIdeo, cytoband = FALSE)
    p + layout_karyogram(c(g.po, g.per), geom = "point", size = 3, 
                         aes(x = midpoint, 
                             y = meth.diff, color = id)) + scale_colour_manual(values = c(hyper.col, 
                                                                                          hypo.col)) + labs(title = title)
    # new alternative commented out
    #autoplot(c(g.po, g.per), layout = "karyogram", geom = "point", size = 0.65, 
    #aes(x = midpoint,y = meth.diff, color = id))  + scale_colour_manual(values = c(hyper.col, 
    #                                                                                        hypo.col)) + labs(title = title)
    
  }
}




require(methylKit)
require(GenomicRanges)
require(ggbio)

# chrom.length
chrom.length = chr.len
myIdeo <-
  GRanges(
    seqnames = names(chrom.length),
    ranges = IRanges(start = 1,
                     width = chrom.length))
  seqlevels(myIdeo) = names(chrom.length)
  seqlengths(myIdeo) = (chrom.length)
  
  
  hypo = getMethylDiff(CpG.Diff, difference = 25, qvalue = 0.01, 
                        type = "hypo")
  print (paste("Number of hypomethylated sites is", nrow(hypo)))
  hyper = getMethylDiff(CpG.Diff, difference = 25, qvalue = 0.01, 
                         type = "hyper")
  print (paste("Number of hypermetyhlated sites is", nrow(hyper)))
  
  g.per = as(hyper, "GRanges")
  seqlevels(g.per, pruning.mode="coarse") <- seqlevels(myIdeo)
  #seqlevels(g.per) = seqlevels(myIdeo)
  seqlengths(g.per)=(chrom.length)
  
  g.po = as(hypo, "GRanges")
  #str(g.po)
  seqlevels(g.po, pruning.mode="coarse") = seqlevels(myIdeo)
  seqlengths(g.po)=(chrom.length)
  
  values(g.po)$id = "Hypomethylated"
  values(g.per)$id = "Hypermethylated"
   #if (circos) {
  
 # seqlevels(myIdeo) <- sub("^chr", "", seqlevels(myIdeo))
  
     pl1 =  ggplot2::ggplot() + layout_circle(myIdeo, geom = "ideo", fill = "black", 
                                  radius = 40, trackWidth = 2) + 
      layout_circle(c(g.po, g.per), geom = "point", 
                           size = 0.2, aes(x = midpoint, 
                                         y = meth.diff, color = id), 
                    radius = 25, trackWidth = 30) +              
      scale_colour_manual(values = c("green", "brown"))+
     layout_circle(myIdeo, geom = "text", aes(label = seqnames), grid = FALSE,  vjust = 0, radius = 55, trackWidth = 5, cex = 9)
    #+ 
      #labs(title = title)
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_CircosPlot_TIvsC.pdf')
pl1 
dev.off()

```


# Karyogram Plot

```{r}
 pl2 =  ggplot() + layout_karyogram(myIdeo)+
    layout_karyogram(c(g.po, g.per), geom = "point", size = 0.2, 
                         aes(x = midpoint, 
                             y = meth.diff, color = id)) +
    scale_colour_manual(values = c(
      "green", "brown")) + labs(title = "Karyogram of CpG", cex = 2)

    # new alternative commented out
    #autoplot(c(g.po, g.per), layout = "karyogram", geom = "point", size = 0.65, 
    #aes(x = midpoint,y = meth.diff, color = id))  + scale_colour_manual(values = c(hyper.col, 
    #                                                                                        hypo.col)) + labs(title = title)

pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_Karyogram_TIvsC.pdf')   
pl2  
dev.off()

```

## Percent Methylation per Chr Bar Plot

```{r, echo = FALSE, percentMethylation, fig.height = 7, fig.width = 6}
par(mar = c(4.1, 4.4, 4.1, 1.9), xaxs="i", yaxs="i")
ChromosomeTable <- diffMethPerChr(CpG.Diff, plot = FALSE, qvalue.cutoff = 0.01, meth.cutoff = 25)
write.csv(ChromosomeTable, "ChromosomeTable.csv")
diff = read.csv("ChromosomeTable.csv")
test_data = data.frame (cbind(diff$diffMeth.per.chr.chr, diff$diffMeth.per.chr.percentage.of.hypermethylated, diff$diffMeth.per.chr.percentage.of.hypomethylated))

test_data$X2 = as.numeric(test_data$X2)

test_data$X3 = as.numeric(test_data$X3)

#str(test_data)
colnames(test_data) = c ("sample", "Hypermethylated", "Hypomethylated")

test_data$sample = factor(test_data$sample, levels = c("chr1",  "chr2",  "chr3",  "chr4",  "chr5",  "chr6",  "chr7",  "chr8",  "chr9",  "chr10",
 "chr11", "chr12", "chr13", "chr14", "chr15", "chr16", "chr17", "chr18", "chr19", "chr20",
"chr21", "chr22", "chr23", "chr24", "chr25", "chr26", "chr27", "chr28", "chr29", "chrX"))

## Instead of theme_minimal() we can use theme_bw()

p1 = ggplot(test_data, aes(Hypermethylated, sample)) +
  geom_col(fill = "green") + theme_bw() + theme(panel.grid.major = element_blank()) +
  theme(panel.border = element_rect(
    fill = NA,
    colour = "black",
    size = 1
  )) +
  scale_x_reverse() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 10),
   # plot.margin = margin(5.5, 0, 5.5, 5.5)
  )

p2 <- ggplot(test_data, aes(Hypomethylated, sample)) +
  geom_col(fill = "brown") + theme_bw() + theme(panel.grid.major = element_blank()) +
  theme(panel.border = element_rect(
    fill = NA,
    colour = "black",
    linewidth = 1
  )) +
  theme(
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_text(size = 10, color = "black"),
    
    axis.ticks.length = unit(0, "pt"),
    axis.text.x = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 10),
   # plot.margin = margin(5.5, 5.5, 5.5, 0.1),
    axis.text.y.left = element_text(margin = margin(0, 5.5, 0, 5.5))
  )

# theme(axis.text=element_text(size=12),
#         axis.title=element_text(size=14,face="bold"))

MPerChr <- ggarrange(p1, p2)


pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_PercentMethylationPerCHR_TIvsC.pdf')
MPerChr
dev.off()

```
# Volcano Plot
```{r}

pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_Volcano_TIvsC.pdf')
CpG.Diff$threshold = with(CpG.Diff, ifelse(abs(CpG.Diff$meth.diff) >= 25 & CpG.Diff$qvalue <= 0.01, sign(CpG.Diff$meth.diff), 1.5))
CpG.Diff$threshold = as.factor(with(CpG.Diff, ifelse(threshold==1, "Hypomethylated", ifelse(threshold==-1, "Hypermethylated", "Not Significant"))))
colors <- c("brown", "green", "black")
plot(CpG.Diff$meth.diff, -log10(CpG.Diff$qvalue), frame.plot=FALSE, pch = 20, main= "4mo CpG Methylation TI vs C", xlab="% Differential Methylation", ylab="-log10(qvalue)", cex.lab = 1.5, cex.axis = 1.5, cex.main = 1.5, col=colors[factor(CpG.Diff$threshold)])
     abline(v=c(-25,25),col="darkgrey",lty=5)
     abline(h=2, col="darkgrey", lty=5)
     box(bty="l")
     legend("top",
            legend = c("Hypomethylated", "Not Significant", "Hypermethylated"), pch=16, cex = 1.2, col = c("brown", "black", "green"))
dev.off()

```

# PCA plots using all CpG sites after filtering but before differential analysis 

```{r}

DMSPositions <- rep(0, times = length(CpG.all$chr)) 
for (i in 1:length(DMSPositions)) {
  DMSPositions[i] <- which(getData(CpG.all)$start[i] == getData(united.CpG)$start)
}  
tail(DMSPositions) 
DMS.Matrix <- united.CpG[DMSPositions,]
# This next function should return 'TRUE', if not, then the vector was not created correctly 
sum((DMS.Matrix$start) == (CpG.all$start)) == length(CpG.all$start)  

sample.id = c("2103.TI", "2106.TI", "2132.C", "2147.C", "2149.C", "2156.C", "2163.C", "2169.TI", "2183.TI", "2186.TI")
# Specify which treatment the samples were from. 1 is the treatment () and 0 is the control ()
## Make sure to reflect the initial order of IDs in numerical order from smallest to largest 
treatment <- c(rep(1, times = 2), rep(0, times = 5), rep(1, times = 3))
allDataPCA = PCASamples(united.CpG, obj.return = TRUE)
plotCustomization <- data.frame(sample = sample.id, treatment) 

pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_TIvsC_AllDataPCA.pdf')
fig.allDataPCA <-
  ordiplot(
    allDataPCA, choices = c(1, 2), type = "none",
    display = "sites", cex = 2, xlab = "",
    ylab = "", xaxt = "n", yaxt = "n") 
points(fig.allDataPCA, "sites", ### Make sure that these repetitions also reflect the order of your ids
       bg = c(rep("grey", times = 2), rep("black", times =5), rep("grey", times = 3)), 
       col = c(rep("black", times =5)), 
       pch = c(rep(21, times = 5), rep(21, times = 5)), , cex = 3) #Add each sample 
ordiellipse(allDataPCA, plotCustomization$treatment, show.groups = "1", 
            col = "grey", lwd = 1.5) #Add confidence ellipse around TI samples
ordiellipse(allDataPCA, plotCustomization$treatment, show.groups = "0", 
            col = "black", lwd = 1.5)

### Make sure to change the PC1 and PC2 values to reflect your own data
axis(side =  1,  labels = TRUE, col = "black", cex.axis = 1.3) #Add x-axis
#mtext(side = 1, text = "PC1 15.13%", line = 3, cex = 1.5) #Add x-axis label

axis(side =  2, labels = TRUE, col = "black", cex.axis = 1.3, las = 1) #Add y-axis
#mtext(side = 2, text = "PC2 12.24%", line = 3, cex = 1.5) #Add y-axis label

legend("topright", 
       legend = c("Control", "TI"), 
       col = c("black", "grey"), 
       pch = c(16, 16), 
       cex = 1.5, bty = "n") #Add legend   

dev.off()
```


```{r}
#Run a PCA analysis on percent methylation for all samples. 
#methylKit uses prcomp to create the PCA matrix
DMSDataPCA <- PCASamples(DMS.Matrix, obj.return = TRUE) 
CpGPCA = PCASamples(united.CpG, obj.return = TRUE)
#Look at summary statistics. 
#The first PC explains 15.13%% of variation, the second PC explains 12.24% of variation
summary(CpGPCA)
summary(DMSDataPCA)
```
# PCA plot for DMS only 

```{r, echo = FALSE}
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_TIvsC_DMSPCA.pdf')
 #Specify inner and outer margins
par(mar = c(5, 5, 1.5, 1.5))
#Use ordiplot to create base biplot. 
fig.DMSDataPCA <- ordiplot(DMSDataPCA, choices = c(1, 2), type = "none", 
                           display = "sites", cex = 0.5, xlab = "", 
                           ylab = "", xaxt = "n", yaxt = "n") 
#Add points
points(fig.DMSDataPCA, "sites", bg = c(rep("grey", times = 2), rep("black", times =5), rep("grey", times =3)), 
       col = c(rep("black", times = 10)), 
       pch = c(rep(21, times = 5), rep(21, times = 5)), cex = 2.5) 
#Add ellipses
ordiellipse(DMSDataPCA, plotCustomization$treatment, show.groups = "1", 
            col = "grey", lwd = 1.5)
ordiellipse(DMSDataPCA, plotCustomization$treatment, show.groups = "0",
            col = "black", lwd = 1.5) 
### Make sure to change the PCA labels to reflect your own data
axis(side =  1, labels = TRUE, col = "black", cex.axis = 1.2) #Add x-axis
mtext(side = 1, text = "PC 1 (22.11%)", line = 3, cex = 1.5) #Add x-axis label

axis(side =  2, labels = TRUE, col = "black", cex.axis = 1.2, las = 1) #Add y-axis
mtext(side = 2, text = "PC 2 (12.48%)", line = 3, cex = 1.5) #Add y-axis label

legend("bottom", 
       pch = c(16, 16), 
       legend = c("Control", "TI"), 
       col = c("black", "grey"), 
       cex = 1.5, bty = "n") 

dev.off()
```


# Heatmap 
```{r, echo = FALSE}
percMethDMS <- percMethylation(DMS.Matrix)
                             
dim(percMethDMS)

percMethDMS <- percMethylation(DMS.Matrix, rowids = TRUE)
head(percMethDMS)
dim(percMethDMS)

# Change the nrow to the value returned from dim(percMethDML)
percMethDMSm = matrix(percMethDMS, nrow = 30470, ncol = 10)
head(percMethDMSm )

colnames(percMethDMSm)<-c("2132.C", "2147.C","2149.C", "2156.C", "2163.C", "2103.TI", "2106.TI", "2169.TI", "2183.TI", "2186.TI")

pdf(file = "/home/jkincade/Methyl/4mo/TI/Plots/4mo_TIvsC_Heatmap.pdf")
par(mar = c(5, 5, 1, 1)) #Specify inner and outer margins
plotColors = colorRampPalette(c("white", "black"),space="rgb")(256)
columns <- data.frame(sample = rep(c("Control", "TI"), c(5,5)))
pheatmap(percMethDMS, color = plotColors, annotation_col = , 
         cluster_rows = TRUE, clustering_distance_rows = "euclidean", treeheight_row = 0, show_rownames = FALSE,
         cluster_cols = TRUE, clustering_distance_cols = "euclidean", treeheight_col = 30, show_colnames = TRUE, angle_col = 45, cex=1.2)
annotation_colors = list(pCO2 = c(Ambient = "black", Treatment = "black"))
annotation_legend = TRUE
annotation_names_col = FALSE
legend = TRUE
dev.off()

```



## Annotating the CpGs

```{r}
## Loading in the annotation data
bov <- readTranscriptFeatures("/home/jkincade/Methyl/bosTau9_RefSeq.txt")



annotated.CpG.hyper <- annotateWithGeneParts(as(CpG.hyper, "GRanges"), bov)
annotated.CpG.hypo <- annotateWithGeneParts(as(CpG.hypo, "GRanges"), bov)
annotated.CpG.all <- annotateWithGeneParts(as(CpG.all, "GRanges"), bov)
```

## Getting the association with transcription start site (TSS) 
```{r}

association.CpG.hyper <- getAssociationWithTSS(annotated.CpG.hyper)
association.CpG.hypo <- getAssociationWithTSS(annotated.CpG.hypo)
association.CpG.all <- getAssociationWithTSS(annotated.CpG.all)

pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_TIvsC_TSS_association.pdf')
hist(association.CpG.all$dist.to.feature[abs(association.CpG.all$dist.to.feature)<=100000],main="distance to nearest TSS",xlab="distance in bp",breaks=50,col="brown4")
dev.off()

#plotTargetAnnotation(annotated.CpG.all, main = "Differential Methylation Annotation")
#getMembers(annotated.CpG.hyper) # whether the DMS are within promoters, introns or exons
#class(association.CpG.all)
#View(association.CpG.all)

genomation::getTargetAnnotationStats(annotated.CpG.all, percentage = TRUE, precedence = TRUE)
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_TIvsC_AnnotationTIeChart.pdf')
genomation::plotTargetAnnotation(annotated.CpG.all, precedence = TRUE, main = "CpG.All.DMC")
dev.off()
genomation::getFeatsWithTargetsStats(annotated.CpG.all, percentage = TRUE)
```


```{r, echo = FALSE}
#OPTIONAL
#This code consttruct the bar plot instead of TIe plot
#pdf (paste0(date,"_DMC_in_regions.pdf"), width = 4, height = 4)

dat = data.frame(Regions = c("exon", "promoter","intergenic" , "intron"), 
                 percent = c(13.09, 11.42, 35.61, 39.88))

dat$Regions = factor(dat$Regions, levels = c("exon", "promoter","intergenic" , "intron"))
DMCregions = ggplot(dat, aes(x = Regions, y = percent, fill = Regions)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  theme_classic() + ylab("Percent of DMC, %") +
  scale_fill_manual(values = c('steelblue2', 'steelblue4', "gold2" , "gold4"))+
  theme(axis.text.y = element_text(size = 16),axis.text.x = element_text(size = 16),
  axis.title.y = element_text(size = 16), axis.title.x = element_text(size = 16))+
  theme(legend.position = "none")

#dev.off()
```

```{r}
dim(association.CpG.hyper)

df.CpG.hyper <- getData(CpG.hyper)
#head(df.CpG.hyper)
dim(df.CpG.hyper)[1]

df.CpG.hyper$REFSEQ <- association.CpG.hyper$feature.name

#tail(df.CpG.hyper)
library(tidyr)
library(clusterProfiler)
df.CpG.hyper <-
  separate(df.CpG.hyper, "REFSEQ", c("REFSEQ", "version"), sep = "\\.")
CpG.hyper.gene.id <-
  bitr(
    df.CpG.hyper$REFSEQ,
    fromType = "REFSEQ",
    toType = c("SYMBOL", "GENENAME", "ENTREZID"),
    OrgDb = "org.Bt.eg.db",
    drop = TRUE
  )
symbol.CpG.hyper <- merge(CpG.hyper.gene.id, df.CpG.hyper, by = "REFSEQ")
#head(symbol.CpG.hyper)
```

```{r}
## CpG Hypo ##
df.CpG.hypo <- getData(CpG.hypo)
#head(df.CpG.hypo)
dim(df.CpG.hypo)
dim(association.CpG.hypo)
df.CpG.hypo$REFSEQ <- association.CpG.hypo$feature.name
df.CpG.hypo <-
  separate(df.CpG.hypo, "REFSEQ", c("REFSEQ", "version"), sep = "\\.")
CpG.hypo.gene.id <-
  bitr(
    df.CpG.hypo$REFSEQ,
    fromType = "REFSEQ",
    toType = c("SYMBOL", "GENENAME", "ENTREZID"),
    OrgDb = "org.Bt.eg.db",
    drop = TRUE
  )
symbol.CpG.hypo <-
  merge(CpG.hypo.gene.id, df.CpG.hypo, by = "REFSEQ")
#symbol.CpG.hypo <- merge(symbol.CpG.hypo, df.CpG.hypo.heatmap [,c(3:10)], by = "end")
#head(symbol.CpG.hypo)
```

```{r}
## CpG All ##
df.CpG.all <- getData(CpG.all)
dim(df.CpG.all)
dim(association.CpG.all)

df.CpG.all$REFSEQ <- association.CpG.all$feature.name
df.CpG.all <-
  separate(df.CpG.all, "REFSEQ", c("REFSEQ", "version"), sep = "\\.")
CpG.all.gene.id <-
  bitr(
    df.CpG.all$REFSEQ,
    fromType = "REFSEQ",
    toType = c("SYMBOL", "GENENAME", "ENTREZID"),
    OrgDb = "org.Bt.eg.db",
    drop = TRUE
  )
symbol.CpG.all <- merge(CpG.all.gene.id, df.CpG.all, by = "REFSEQ")
#symbol.CpG.all <- merge(symbol.CpG.all, df.CpG.all.heatmap [,c(3:10)], by = "end")
#head(symbol.CpG.all)

```

```{r}
symbol.CpG.hyper$context <- "CpG"
symbol.CpG.hyper$state <- "Hyper"
symbol.CpG.hypo$context <- "CpG"
symbol.CpG.hypo$state <- "Hypo"

merged <- rbind(symbol.CpG.hyper, symbol.CpG.hypo)
```


```{r}
date = format(Sys.Date(), '%m-%d-%Y')
date
write.csv(
  symbol.CpG.hypo,
  file = paste(folder, "4mo-TIvsC-CpG-DMS-Hypo_", date, ".csv", sep = ""),
  row.names = FALSE
)
write.csv(
  symbol.CpG.hyper,
  file = paste(folder, "4mo-TIvsC-CpG-DMS-Hyper_", date, ".csv", sep = ""),
  row.names = FALSE
)
write.csv(
  symbol.CpG.all,
  file = paste(folder, "4mo-TIvsC-CpG-DMS-All_", date, ".csv", sep = ""),
  row.names = FALSE
)
```


```{r}
length(unique(symbol.CpG.hyper$GENENAME))
length(unique(symbol.CpG.hypo$GENENAME))
```
Venn Diagram here have one drawback: it calculates unique sites (but one gene can be methylated in multiple sites). Therefore need to think differently how to present the data.

```{r}
# library(VennDiagram)
# 
# set1hyper = read.csv("/home/dmurtazi/hanahm1/DataFilesHansen/Islands/CpG-hyper-noSNP_01-26-2024.csv")
# 
# set2hypo = read.csv("/home/dmurtazi/hanahm1/DataFilesHansen/Islands/CpG-Hypo-noSNP_01-26-2024.csv")
# 
# set1 = set1hyper[, 1]
# length(set1)
# length(unique(set1))
# n_occur <- data.frame(table(set1hyper$REFSEQ)) # create table with occurences
# n_occur = n_occur[n_occur$Freq > 1,] # how many specific REFSEq occurs in the data set
# 
# 
# set2 = set2hypo[, 1]
# length(set2)
# length(unique(set2))
# 
# myCol = c("orange", "lightblue")
# 
# # Chart
# venn.diagram(
#         x = list(set1, set2),
#         category.names = c("Hyper" , "Hypo" ),
#          filename = paste(date, '_venn_diagramm.png'), scaled = TRUE,
#          output=TRUE,
#         
#         # Output features
#         # imagetype="png" ,
#         # height = 780 ,
#         # width = 780 ,
#         # resolution = 600,
#         # compression = "lzw",
#         
#         # Circles
#         lwd = 2,
#         lty = 'blank',
#         fill = c("orange", "lightblue"),
#         #fill = myCol,
#         
#         # Numbers
#         cex = .6,
#         fontface = "bold",
#         fontfamily = "sans"
# )
```


```{r}
#####################################################################################
## GO and KEGG ##

CpG.all.GO <-
  groupGO(
    gene = symbol.CpG.all$ENTREZID,
    "org.Bt.eg.db",
    ont = "BP",
    level = 3,
    readable = TRUE
  )
CpG.hyper.GO <-
  groupGO(
    gene = symbol.CpG.hyper$ENTREZID,
    "org.Bt.eg.db",
    ont = "BP",
    level = 3,
    readable = TRUE
  )
CpG.hypo.GO <-
  groupGO(
    gene = symbol.CpG.hypo$ENTREZID,
    "org.Bt.eg.db",
    ont = "BP",
    level = 3,
    readable = TRUE
  )
```

```{r}
head(CpG.all.GO)
```




```{r}
#print ("GETTING BARPLOTS WITH PATHWAYS")

x = CpG.all.GO[order(CpG.all.GO$Count, decreasing = TRUE) ]
#head(x, 30)
#x$GeneRatio = as.numeric(x$GeneRatio)
#head(head(x, 30))
#CpG.hyper.GO[order(CpG.hyper.GO$Count), ]
#CpG.hypo.GO[order(CpG.hypo.GO$Count), ]

pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_TIvsC_DMS_KEGG-GO.pdf')
par(mar = c(4.1, 4.4, 4.1, 1.9), xaxs="i", yaxs="i")

barplot(CpG.all.GO, drop = TRUE, showCategory = 10, xlab = "Gene Count", title = "CpG.all")
barplot(CpG.hyper.GO, drop = TRUE, showCategory = 10, xlab = "Gene Count", title = "CpG.hyper")
barplot(CpG.hypo.GO, drop = TRUE, showCategory = 10, xlab = "Gene Count", title = "CpG.hypo")

dev.off()

#####################################################################################

```

```{r}
# ego3 <- gseGO(geneList     = geneList,
#               OrgDb        = org.Hs.eg.db,
#               ont          = "CC",
#               minGSSize    = 100,
#               maxGSSize    = 500,
#               pvalueCutoff = 0.05,
#               verbose      = FALSE)
# 
# 
# goplot(CpG.all.GO)
```

We can also summarize methylation information over a set of defined regions such as promoters or CpG islands. The function below summarizes the methylation information over a given set of promoter regions and outputs a `methylRaw` or `methylRawList` object depending on the input. We are using the output of `genomation` functions used above to provide the locations of promoters. For regional summary functions, we need to provide regions of interest as GRanges object.




# CpG islands 

```{r}
# read the shores and flanking regions and name the flanks as shores 
# and CpG islands as CpGi
bov_isl = readFeatureFlank(
  "/home/jkincade/Methyl/bosTau9.CpG.Islands.txt",
  feature.flank.name = c("CpGi", "shores")
)
# convert methylDiff object to GRanges and annotate
# > All_ga=annotateWithGeneParts(as(All_df, "GRanges"), hg.19)
CpG.all_isl <-
  annotateWithFeatureFlank(
    as(CpG.all, "GRanges"),
    bov_isl$CpGi,
    bov_isl$shores,
    feature.name = "CpGi",
    flank.name = "shores"
  )

CpG.all_isl 

```


```{r}
#genomation::getMembers(CpG.all_isl )
genomation::getTargetAnnotationStats(CpG.all_isl )
```


```{r}
myobj_islands <- regionCounts(CpG.Read, bov_isl$CpGi)
# Filter the summarized counts by coverage
myobj_islands_filt <- filterByCoverage(myobj_islands,
                                       lo.count=10,
                                       lo.perc=NULL,
                                       hi.count=NULL,
                                       hi.perc=99.9)
# Perform simple normalization
myobj_islands_filt_norm <- normalizeCoverage(myobj_islands_filt, method = "median")
# Merge the samples again
meth_islands <- methylKit::unite(myobj_islands_filt_norm, destrand=FALSE)
```
```{r}
#Now differential methylation is performed as for the single CpGs
#Test for the differential Methylation
diff.islands = calculateDiffMeth(meth_islands, mc.cores = 30)
```
# Removing 'chrUNs' (chromosome unknown)

```{r}
rowlist <- list()
L <- 1
for(i in 1:nrow(diff.islands)){
  chr <- diff.islands$chr [i]
  if(grepl("chrUn",chr)){
    rowlist [[L]] <- i
    L <- L + 1
  }
}
#rowlist
rw.length <- length(rowlist)
rw.length.minus <- rowlist[[1]] - 1
rowlist[[rw.length]]
diff.islands <- diff.islands[-c(rowlist[[1]]:rowlist[[rw.length]]),]
```




Use diff = 15


```{r, echo = FALSE}
#get differentially methylated CpG Islands
diff.islands.all =  getMethylDiff(diff.islands, difference = 15, qvalue = 0.01)
diff.islands.hyper =  getMethylDiff(diff.islands, difference = 15, qvalue = 0.01, type = "hyper")
diff.islands.hypo =  getMethylDiff(diff.islands, difference = 15, qvalue = 0.01, type = "hypo")

print (paste("Number of sites associated with islands is", dim(diff.islands.all)[1]))
print (paste("Number of  hyperMethylated sites associated with islands is", dim(diff.islands.hyper)[1]))
print (paste("Number of  hypoMethylated sites associated with islands is", dim(diff.islands.hypo)[1]))

```


```{r, echo =FALSE}
a2=  dim(getMethylDiff(diff.islands, difference = 10, qvalue = 0.01))[1]
b2 = dim(getMethylDiff(diff.islands, difference = 15, qvalue = 0.01))[1]
c2 = dim(getMethylDiff(diff.islands, difference = 20, qvalue = 0.01))[1]
d2 = dim(getMethylDiff(diff.islands, difference = 25, qvalue = 0.01))[1]

print (paste("Number of sites associated with islands with diff 10 is", a2))
print (paste("Number of sites associated with islands with diff 15 is", b2))
print (paste("Number of sites associated with islands with diff 20 is", c2))
print (paste("Number of sites associated with islands with diff 25 is", d2))
       
```




```{r, echo = FALSE}
diff = c(10,15, 20, 25)
DMR_isl = c(a2, b2, c2, d2)
df_DMR_isl = data.frame(diff, DMR_isl)


pl_Diff_CpG = ggplot(df_DMR_isl, aes(x = diff, y = DMR_isl)) +
  geom_point(size = 3) +
  
    geom_line(col = "blue")+
 
  ylab("Number of DMR (CpGi)") + xlab("Parameters of diff")+
  geom_vline(xintercept = 25, linetype = "dashed", color = "black", size = 0.5)+
  geom_hline(yintercept = 8, linetype = "dashed", color = "black", size = 0.5)+
  geom_vline(xintercept = 15, linetype = "dashed", color = "red", size = 0.5)+
  geom_hline(yintercept = 89, linetype = "dashed", color = "red", size = 0.5)+
  theme_classic()+

theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))

pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/TIvsC-CpG-islands-WandWO-overdispcor.pdf')
pl_Diff_CpG
dev.off()
```

Choose difference 15 (red dotted line)

#Using Genomation

```{r}
myDiff_isl_annot <- annotateWithGeneParts(as((diff.islands.all), "GRanges"), bov)
myDiff_isl_hyper_annot <- annotateWithGeneParts(as((diff.islands.hyper), "GRanges"), bov)
myDiff_isl_hypo_annot <- annotateWithGeneParts(as((diff.islands.hypo), "GRanges"), bov)
# View the distance to the nearest Transcription Start Site; 
#the target.row column indicates the row number in myDiff_islands_25p
association.CpGi.all = getAssociationWithTSS(myDiff_isl_annot )
association.CpGi.hyper = getAssociationWithTSS(myDiff_isl_hyper_annot)
association.CpGi.hypo = getAssociationWithTSS(myDiff_isl_hypo_annot)

```



```{r}

#dim(association.CpGp.all)
df.CpGi.hyper <- getData(diff.islands.hyper)

dim(df.CpGi.hyper)
df.CpGi.hypo <- getData(diff.islands.hypo)

dim(df.CpGi.hypo)
df.CpGi.all <- getData(diff.islands.all)

dim(df.CpGi.all)

```

```{r}
df.CpGi.hyper$REFSEQ <- association.CpGi.hyper$feature.name

df.CpGi.hyper <-
  separate(df.CpGi.hyper, "REFSEQ", c("REFSEQ", "version"), sep = "\\.")
CpGi.hyper.gene.id <-
  bitr(
    df.CpGi.hyper$REFSEQ,
    fromType = "REFSEQ",
    toType = c("SYMBOL", "GENENAME", "ENTREZID"),
    OrgDb = "org.Bt.eg.db",
    drop = TRUE
  )
symbol.CpGi.hyper <-
  merge(CpGi.hyper.gene.id, df.CpGi.hyper, by = "REFSEQ")
#head(symbol.CpGi.hyper, 2)

df.CpGi.hypo$REFSEQ <- association.CpGi.hypo$feature.name

df.CpGi.hypo <-
  separate(df.CpGi.hypo, "REFSEQ", c("REFSEQ", "version"), sep = "\\.")
CpGi.hypo.gene.id <-
  bitr(
    df.CpGi.hypo$REFSEQ,
    fromType = "REFSEQ",
    toType = c("SYMBOL", "GENENAME", "ENTREZID"),
    OrgDb = "org.Bt.eg.db",
    drop = TRUE
  )
symbol.CpGi.hypo <-
  merge(CpGi.hypo.gene.id, df.CpGi.hypo, by = "REFSEQ")
#head(symbol.CpGi.hypo, 2)


df.CpGi.all$REFSEQ <- association.CpGi.all$feature.name

df.CpGi.all <-
  separate(df.CpGi.all, "REFSEQ", c("REFSEQ", "version"), sep = "\\.")
CpGi.all.gene.id <-
  bitr(
    df.CpGi.all$REFSEQ,
    fromType = "REFSEQ",
    toType = c("SYMBOL", "GENENAME", "ENTREZID"),
    OrgDb = "org.Bt.eg.db",
    drop = TRUE
  )
symbol.CpGi.all <-
  merge(CpGi.all.gene.id, df.CpGi.all, by = "REFSEQ")
#head(symbol.CpGi.all, 2)
```

```{r}
write.csv(merged, file = paste(folder, "merged-DMRs", date, ".csv", sep = ""), row.names = FALSE)
date = format(Sys.Date(), '%m-%d-%Y')
#date
write.csv(
  symbol.CpGi.hypo,
  file = paste(folder, "CpGi-Hypo_", date, ".csv", sep = ""),
  row.names = FALSE
)
write.csv(
  symbol.CpGi.hyper,
  file = paste(folder, "CpGi-hyper_", date, ".csv", sep = ""),
  row.names = FALSE
)
write.csv(
  symbol.CpGi.all,
  file = paste(folder, "CpGi-all_", date, ".csv", sep = ""),
  row.names = FALSE
)
```


```{r}
symbol.CpGi.hyper$context <- "CpG"
symbol.CpGi.hyper$state <- "Hyper"
symbol.CpGi.hypo$context <- "CpG"
symbol.CpGi.hypo$state <- "Hypo"

merged <- rbind(symbol.CpGi.hyper, symbol.CpGi.hypo)

merged$Pred.Expr <- merged$meth.diff * -1
```


```{r}
CpGi.all.GO <-
  groupGO(
    gene = symbol.CpGi.all$ENTREZID,
    "org.Bt.eg.db",
    ont = "BP",
    level = 3,
    readable = TRUE
  )
CpGi.hyper.GO <-
  groupGO(
    gene = symbol.CpGi.hyper$ENTREZID,
    "org.Bt.eg.db",
    ont = "BP",
    level = 3,
    readable = TRUE
  )
CpGi.hypo.GO <-
  groupGO(
    gene = symbol.CpGi.hypo$ENTREZID,
    "org.Bt.eg.db",
    ont = "BP",
    level = 3,
    readable = TRUE
  )

Islands_all <- as.data.frame(CpGi.all.GO)

Islands <-  Islands_all %>%
  filter(Count !=0)
#Islands 


write.csv(
  Islands,
  file = paste(
    folder,
    "Islands_GO",
    date,
    ".csv",
    quote = FALSE,
    sep = ""
  ),
  row.names = FALSE
)
```



# Promoters

```{r}
promoters = regionCounts(CpG.Read, bov$promoters)
#head(promoters[[1]])
```



```{r}
promoters_filt <- filterByCoverage(promoters,
                                       lo.count=10,
                                       lo.perc=NULL,
                                       hi.count=NULL,
                                       hi.perc=99.9)
# Perform simple normalization
promoters_filt_norm <- normalizeCoverage(promoters_filt, method = "median")
# Merge the samples again
meth_promoters <- methylKit::unite(promoters_filt_norm, destrand=FALSE)
```


```{r}
# Test for differential methylation... This might take a few minutes.
myDiff_promoters <- calculateDiffMeth(meth_promoters, mc.cores=30)
# Rank by significance
myDiff_promoters <- myDiff_promoters[order(myDiff_promoters$qvalue),]
```

```{r}
dim(getMethylDiff(myDiff_promoters, difference = 15, qvalue = 0.01))[1]
```


```{r}
# get all differentially methylated promoters
myDiff_promoters_all <-
  getMethylDiff(myDiff_promoters, difference = 15, qvalue = 0.01)
myDiff_promoters_hyper <-
  getMethylDiff(
    myDiff_promoters,
    difference = 15,
    qvalue = 0.01,
    type = "hyper"
  )
myDiff_promoters_hypo <-
  getMethylDiff(
    myDiff_promoters,
    difference = 15,
    qvalue = 0.01,
    type = "hypo"
  )
```


```{r, echo = FALSE}
a3 = dim(getMethylDiff(myDiff_promoters, difference = 10, qvalue = 0.01))[1]
b3 = dim(getMethylDiff(myDiff_promoters, difference = 15, qvalue = 0.01))[1]
c3 = dim(getMethylDiff(myDiff_promoters, difference = 20, qvalue = 0.01))[1]
d3 = dim(getMethylDiff(myDiff_promoters, difference = 25, qvalue = 0.01))[1]

print (paste("Number of sites associated with promoters with diff 10 is", a3))
print (paste("Number of sites associated with promoters with diff 15 is", b3))
print (paste("Number of sites associated with promoters with diff 20 is", c3))
print (paste("Number of sites associated with promoters with diff 25 is", d3))

```



```{r, echo = FALSE}
diff = c(10,15, 20, 25)
DMR_prm = c(a3, b3, c3, d3)
df_DMR_prm = data.frame(diff, DMR_prm)
#df_DMR_prm
```

```{r, echo =FALSE}

pl_diff_prom = ggplot(df_DMR_isl, aes(x = diff, y = DMR_prm)) +
  geom_point(size = 3) +
  
    geom_line(col = "blue")+
 
  ylab("Number of DMR (promoters)") + xlab("Parameters of diff")+
  geom_vline(xintercept = 25, linetype = "dashed", color = "black", size = 0.5)+
  geom_hline(yintercept = 8, linetype = "dashed", color = "black", size = 0.5)+
  geom_vline(xintercept = 15, linetype = "dashed", color = "red", size = 0.5)+
  geom_hline(yintercept = 84, linetype = "dashed", color = "red", size = 0.5)+
  theme_classic()+

theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))

pdf(file= '/home/jkincade/Methyl/4mo/TI/Plots/TIvsC-Promoters-WandWO-overdispcor.pdf')
pl_diff_prom
dev.off()
```

Choose difference 15 (red dotted line)

#Using Genomation

```{r}
myDiff_promoters_annot <- annotateWithGeneParts(as((myDiff_promoters_all), "GRanges"), bov)
myDiff_promoters_hyper_annot <- annotateWithGeneParts(as((myDiff_promoters_hyper), "GRanges"), bov)
myDiff_promoters_hypo_annot <- annotateWithGeneParts(as((myDiff_promoters_hypo), "GRanges"), bov)
# View the distance to the nearest Transcription Start Site; 
#the target.row column indicates the row number in myDiff_islands_25p
association.CpGp.all = getAssociationWithTSS(myDiff_promoters_annot)
association.CpGp.hyper = getAssociationWithTSS(myDiff_promoters_hyper_annot)
association.CpGp.hypo = getAssociationWithTSS(myDiff_promoters_hypo_annot)

```


```{r}
#dim(association.CpGp.all)
df.CpGp.hyper <- getData(myDiff_promoters_hyper)
#head(df.CpGp.hyper)
dim(df.CpGp.hyper)
```


```{r}
df.CpGp.hyper$REFSEQ <- association.CpGp.hyper$feature.name
#View(df.CpGi.hyper)
dim (df.CpGp.hyper)
#tail(df.CpGp.hyper)
df.CpGp.hyper <-
  separate(df.CpGp.hyper, "REFSEQ", c("REFSEQ", "version"), sep = "\\.")
CpGp.hyper.gene.id <-
  bitr(
    df.CpGp.hyper$REFSEQ,
    fromType = "REFSEQ",
    toType = c("SYMBOL", "GENENAME", "ENTREZID"),
    OrgDb = "org.Bt.eg.db",
    drop = TRUE
  )
symbol.CpGp.hyper <-
  merge(CpGp.hyper.gene.id, df.CpGp.hyper, by = "REFSEQ")
#head(symbol.CpGp.hyper)
```

```{r}
#View(symbol.CpGp.hyper)

## CpGi Hypo ##
df.CpGp.hypo <- getData(myDiff_promoters_hypo)
#head(df.CpGp.hypo)
dim(df.CpGp.hypo)
#df.CpGp.hypo

dim(association.CpGp.hypo)
df.CpGp.hypo$REFSEQ <- association.CpGp.hypo$feature.name
df.CpGp.hypo <-
  separate(df.CpGp.hypo, "REFSEQ", c("REFSEQ", "version"), sep = "\\.")
CpGp.hypo.gene.id <-
  bitr(
    df.CpGp.hypo$REFSEQ,
    fromType = "REFSEQ",
    toType = c("SYMBOL", "GENENAME", "ENTREZID"),
    OrgDb = "org.Bt.eg.db",
    drop = TRUE
  )
symbol.CpGp.hypo <-
  merge(CpGp.hypo.gene.id, df.CpGp.hypo, by = "REFSEQ")
#symbol.CpG.hypo <- merge(symbol.CpG.hypo, df.CpG.hypo.heatmap [,c(3:10)], by = "end")

#head(symbol.CpGp.hypo)
```

```{r}
## CpGi All ##
df.CpGp.all <- getData(myDiff_promoters_all)
dim(df.CpGp.all)

df.CpGp.all$REFSEQ <- association.CpGp.all$feature.name
df.CpGp.all <-
  separate(df.CpGp.all, "REFSEQ", c("REFSEQ", "version"), sep = "\\.")
CpGp.all.gene.id <-
  bitr(
    df.CpGp.all$REFSEQ,
    fromType = "REFSEQ",
    toType = c("SYMBOL", "GENENAME", "ENTREZID"),
    OrgDb = "org.Bt.eg.db",
    drop = TRUE
  )
symbol.CpGp.all <-
  merge(CpGp.all.gene.id, df.CpGp.all, by = "REFSEQ")
#symbol.CpG.all <- merge(symbol.CpG.all, df.CpG.all.heatmap [,c(3:10)], by = "end")
#head(symbol.CpGp.all)
```


```{r}
symbol.CpGp.hyper$context <- "CpG"
symbol.CpGp.hyper$state <- "Hyper"
symbol.CpGp.hypo$context <- "CpG"
symbol.CpGp.hypo$state <- "Hypo"

merged <- rbind(symbol.CpGp.hyper, symbol.CpGp.hypo)

merged$Pred.Expr <- merged$meth.diff * -1
```


```{r}
#write.csv(merged, file = paste(folder, "merged-DMRs", date, ".csv", sep = ""), row.names = FALSE)
date = format(Sys.Date(), '%m-%d-%Y')
write.csv(
  symbol.CpGp.hypo,
  file = paste(folder, "CpGp-Hypo_", date, ".csv", sep = ""),
  row.names = FALSE
)
write.csv(
  symbol.CpGp.hyper,
  file = paste(folder, "CpGp-hyper_", date, ".csv", sep = ""),
  row.names = FALSE
)
write.csv(
  symbol.CpGp.all,
  file = paste(folder, "CpGp-all_", date, ".csv", sep = ""),
  row.names = FALSE
)
```

```{r}
## GO and KEGG ##

CpGp.all.GO <-
  groupGO(
    gene = symbol.CpGp.all$ENTREZID,
    "org.Bt.eg.db",
    ont = "BP",
    level = 3,
    readable = TRUE
  )
CpGp.hyper.GO <-
  groupGO(
    gene = symbol.CpGp.hyper$ENTREZID,
    "org.Bt.eg.db",
    ont = "BP",
    level = 3,
    readable = TRUE
  )
CpGp.hypo.GO <-
  groupGO(
    gene = symbol.CpGp.hypo$ENTREZID,
    "org.Bt.eg.db",
    ont = "BP",
    level = 3,
    readable = TRUE
  )

Promoter_all <- as.data.frame(CpGp.all.GO)

Promoter_GO <-  Promoter_all %>%
  filter(Count !=0)
Promoter_GO 

write.csv(
  Promoter_GO,
  file = paste(folder, "Promoter_GO", date, ".csv", sep = ""),
  row.names = FALSE
)
```

## GETTING BARTI/Plots WITH PATHWAYS

```{r}
#pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_IslandBarplotPathways TIvsC.pdf')
par(mar = c(4.1, 4.4, 4.1, 1.9), xaxs="i", yaxs="i")

barplot(
  CpGp.all.GO,
  drop = TRUE,
  showCategory = 10,
  xlab = "Gene Count",
  title = "CpG.all"
)
# barplot(CpGp.hyper.GO, drop = TRUE, showCategory = 10, xlab = "Gene Count", title = "CpG.hyper")
barplot(
  CpGi.hypo.GO,
  drop = TRUE,
  showCategory = 10,
  xlab = "Gene Count",
  title = "CpG.hypo"
)
#dev.off()
```

# Tiling window analysis

### 1000 bp windows

```{r}

#takes about 10 min
tiles1000 <-
  tileMethylCounts(
    CpG.Read,
    win.size = 1000,
    step.size = 1000,
    cov.bases = 10,
    num.cores = 30
  ) #Add up C and T counts from each covered cytosine and return total C and T count for each tile. Perform this analysis with 100 bp sized windows, and slide 100 bp for each new window
head(tiles1000[[10]]) #Confirm object creation. [[10]] only shows tile100 head for the 10th sample processed.
```

```{r}
meth.tiles = methylKit::unite(tiles1000, destrand=TRUE) 
```

```{r}
myDiff.tiles = calculateDiffMeth(meth.tiles, num.cores=30)
```
# Removing 'chrUNs' (chromosome unknown)

```{r}
rowlist <- list()
L <- 1
for(i in 1:nrow(myDiff.tiles)){
  chr <- myDiff.tiles$chr [i]
  if(grepl("chrUn",chr)){
    rowlist [[L]] <- i
    L <- L + 1
  }
}
#rowlist
rw.length <- length(rowlist)
rw.length.minus <- rowlist[[1]] - 1
rowlist[[rw.length]]
myDiff.tiles <- myDiff.tiles[-c(rowlist[[1]]:rowlist[[rw.length]]),]
```

```{r}
dim(getMethylDiff(myDiff.tiles,qvalue=0.01,difference=10,type="all"))[1]
dim(getMethylDiff(myDiff.tiles,qvalue=0.01,difference=15,type="all"))[1]
dim(getMethylDiff(myDiff.tiles,qvalue=0.01,difference=20,type="all"))[1]
dim(getMethylDiff(myDiff.tiles,qvalue=0.01,difference=25,type="all"))[1]
```



```{r}
myDiff.tiles.hyper = getMethylDiff(myDiff.tiles,qvalue=0.01,difference=15,type="hyper")
bedgraph(myDiff.tiles.hyper, file.name = "hyper.DMRTiles.bedGraph", col.name = "meth.diff")

myDiff.tiles.hypo = getMethylDiff(myDiff.tiles,qvalue=0.01,difference=15,type="hypo")
bedgraph(myDiff.tiles.hypo, file.name = "hypo.DMRtiles.bedGraph", col.name = "meth.diff")

myDiff.tiles.all = getMethylDiff(myDiff.tiles,qvalue=0.01,difference=15,type="all")
bedgraph(myDiff.tiles.all, file.name = "all.DMRtiles.bedGraph", col.name = "meth.diff")
```

Using Genomation

```{r}
myDiff_tiles_annot <- annotateWithGeneParts(as((myDiff.tiles.all), "GRanges"), bov)
myDiff_tiles_hyper_annot <- annotateWithGeneParts(as((myDiff.tiles.hyper), "GRanges"), bov)
myDiff_tiles_hypo_annot <- annotateWithGeneParts(as((myDiff.tiles.hypo), "GRanges"), bov)
# View the distance to the nearest Transcription Start Site; 
#the target.row column indicates the row number in myDiff_islands_25p
association.CpGt.all = getAssociationWithTSS(myDiff_tiles_annot)
association.CpGt.hyper = getAssociationWithTSS(myDiff_tiles_hyper_annot)
association.CpGt.hypo = getAssociationWithTSS(myDiff_tiles_hypo_annot)

```

```{r}
#dim(association.CpGp.all)
df.CpGt.hyper <- getData(myDiff.tiles.hyper)
#head(df.CpGp.hyper)
dim(df.CpGt.hyper)

df.CpGt.hypo <- getData(myDiff.tiles.hypo)
#head(df.CpGp.hypo)
dim(df.CpGt.hypo)

df.CpGt.all <- getData(myDiff.tiles.all)
#head(df.CpGp.hyper)
dim(df.CpGt.all)
```

```{R}
df.CpGt.hyper$REFSEQ <- association.CpGt.hyper$feature.name
df.CpGt.hyper <-
  separate(df.CpGt.hyper, "REFSEQ", c("REFSEQ", "version"), sep = "\\.")
CpGt.hyper.gene.id <-
  bitr(
    df.CpGt.hyper$REFSEQ,
    fromType = "REFSEQ",
    toType = c("SYMBOL", "GENENAME", "ENTREZID"),
    OrgDb = "org.Bt.eg.db",
    drop = TRUE
  )
symbol.CpGt.hyper <-
  merge(CpGt.hyper.gene.id, df.CpGt.hyper, by = "REFSEQ")
#head(symbol.CpGp.hyper)


df.CpGt.hypo$REFSEQ <- association.CpGt.hypo$feature.name
df.CpGt.hypo <-
  separate(df.CpGt.hypo, "REFSEQ", c("REFSEQ", "version"), sep = "\\.")
CpGt.hypo.gene.id <-
  bitr(
    df.CpGt.hypo$REFSEQ,
    fromType = "REFSEQ",
    toType = c("SYMBOL", "GENENAME", "ENTREZID"),
    OrgDb = "org.Bt.eg.db",
    drop = TRUE
  )
symbol.CpGt.hypo <-
  merge(CpGt.hypo.gene.id, df.CpGt.hypo, by = "REFSEQ")

df.CpGt.all$REFSEQ <- association.CpGt.all$feature.name
df.CpGt.all <-
  separate(df.CpGt.all, "REFSEQ", c("REFSEQ", "version"), sep = "\\.")
CpGt.all.gene.id <-
  bitr(
    df.CpGt.all$REFSEQ,
    fromType = "REFSEQ",
    toType = c("SYMBOL", "GENENAME", "ENTREZID"),
    OrgDb = "org.Bt.eg.db",
    drop = TRUE
  )
symbol.CpGt.all <-
  merge(CpGt.all.gene.id, df.CpGt.all, by = "REFSEQ")
```

```{r}
symbol.CpGt.hyper$context <- "CpG"
symbol.CpGt.hyper$state <- "Hyper"
symbol.CpGt.hypo$context <- "CpG"
symbol.CpGt.hypo$state <- "Hypo"

merged <- rbind(symbol.CpGt.hyper, symbol.CpGt.hypo)

merged$Pred.Expr <- merged$meth.diff * -1
```


```{r}
# write.csv(merged, file = paste(folder, "merged-DMRs", date, ".csv", sep = ""), row.names = FALSE)
date = format(Sys.Date(), '%m-%d-%Y')
write.csv(
  symbol.CpGt.hypo,
  file = paste(folder, "CpGt-Hypo_", date, ".csv", sep = ""),
  row.names = FALSE
)
write.csv(
  symbol.CpGt.hyper,
  file = paste(folder, "CpGt-hyper_", date, ".csv", sep = ""),
  row.names = FALSE
)
write.csv(
  symbol.CpGt.all,
  file = paste(folder, "CpGt-all_", date, ".csv", sep = ""),
  row.names = FALSE
)
```




```{r}
## GO and KEGG ##

CpGt.all.GO <-
  groupGO(
    gene = symbol.CpGt.all$ENTREZID,
    "org.Bt.eg.db",
    ont = "BP",
    level = 3,
    readable = TRUE
  )
CpGt.hyper.GO <-
  groupGO(
    gene = symbol.CpGt.hyper$ENTREZID,
    "org.Bt.eg.db",
    ont = "BP",
    level = 3,
    readable = TRUE
  )
CpGt.hypo.GO <-
  groupGO(
    gene = symbol.CpGt.hypo$ENTREZID,
    "org.Bt.eg.db",
    ont = "BP",
    level = 3,
    readable = TRUE
  )

#Write to the file
CpGGOCategories <- as.data.frame(CpGt.all.GO)
library(dplyr)

Tiles1000_GO <-  CpGGOCategories %>%
  filter(Count != 0)
Tiles1000_GO
write.csv(
  Tiles1000_GO,
  file = paste(folder, "Tiles1000_GO", date, ".csv", sep = ""),
  row.names = FALSE
)
```

## GETTING BARPLOTS WITH PATHWAYS

```{r}
pdf(file = '/home/jkincade/Methyl/4mo/TI/Plots/4mo_TIvsC_Pathways_tiles500.pdf')
#par(mar = c(4.1, 4.4, 4.1, 1.9), xaxs="i", yaxs="i")
par(mfrow = c(1, 3))


barplot(
  CpGt.all.GO,
  drop = TRUE,
  showCategory = 10,
  xlab = "Gene Count",
  title = "CpG.all"
)
barplot(
  CpGt.hyper.GO,
  drop = TRUE,
  showCategory = 10,
  xlab = "Gene Count",
  title = "CpG.hyper"
)
barplot(
  CpGt.hypo.GO,
  drop = TRUE,
  showCategory = 10,
  xlab = "Gene Count",
  title = "CpG.hypo"
)
dev.off()
```

